Community-Conditioned Language Models

### 2021-05-14 paper/paper.tex
typography + minor fixes and questions

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

@@

Second, we want to find out how, and how much, CCLMs are attuned to
the specificity of communities. We test whether 1.\ the linguistic
models are correlated with a model of communities based only on
co-occurrence of users between communities and 2.\ if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language becomes and the easier it is to identify.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{floats/confusion_lstm-3-1}
\caption{
  Log-probability LMCC confusion matrices for the LSTM model with $l_c=1$. 
  Communities are sorted by indiscernability.}
\label{fig:confusion}
\end{figure}

@@

Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
solution is
\[d(L,S) = \mathsf {Tr} Σ\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). Each
singular \(\sigma_i\) (elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

Further, we can apply a permutation to \(U\),\(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). That
is, the largest singular value \(σ_0\) then corresponds to the principal
directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
principal direction, etc.

The \(d(L,S)\) metric range from \(0\) (corresponding to perfect
independence) to \(n\) (corresponding to perfect correlation, obtained
for example if \(L=S\)), where $n=510$ is the number of communities
considered.

To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
We have computed experimentally \(d(L,S)\) for 
\(L\) and \(S\) chosen as random 16 dimensions. We find
a mean of TODO and a standard deviation of TODO.

@@

(Another idea may be to correlate the cosine similarity between
\((S_i,S_j)\) and \((L_i,L_j)\). But this would yield one number for
each pair of communities (510 * 509 / 2): too much data to deal with,
and it is hard to get a sense of the statistical significance of the
resulting correlation coefficient.)

>>>>

% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage{newunicodechar}
\newunicodechar{Σ}{\ensuremath{\Sigma}} % 
\newunicodechar{σ}{\ensuremath{\mathnormal{\sigma}}} % 
\newunicodechar{μ}{\ensuremath{\mathnormal{\mu}}} % 
\newunicodechar{×}{\ensuremath{\times}} %

@@

Second, we want to find out how, and how much, CCLMs are attuned to
the specificity of communities. We test whether 1.\ the linguistic
models are correlated with a model of communities based only on
co-occurrence of users between communities and 2.\ if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, \jp{easiest it is to model computationally ...}
% the more specific the language becomes and
the easier it is to identify.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{floats/confusion_lstm-3-1}
\caption{
  Log-probability LMCC confusion matrix for the LSTM model with $l_c=1$. 
  Communities are sorted by indiscernability. (The first communities predominantly use a language other than English, and are thus very discernable.)}
\label{fig:confusion}
\end{figure}

@@

Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
solution is
\[d(L,S) = \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). Each
singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

Further, we can apply a permutation to \(U\), \(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). Doing so, the largest singular value \(σ_0\) corresponds to the principal
directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
principal direction, etc.

The \(d(L,S)\) metric ranges from \(0\) (corresponding to perfect
independence) to \(n\) (corresponding to perfect correlation, obtained
for example if \(L=S\)), where $n=510$ is the number of communities
considered.

To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
We have computed experimentally \(d(L,S)\) for 
\(L\) and \(S\) chosen as random 16 dimensions. We find
a mean \(μ_d\)=TODO and a standard deviation \(σ_d\)=TODO.\jp{Unfortunate overloading}

@@

(Another idea may be to correlate the cosine similarity between
\((S_i,S_j)\) and \((L_i,L_j)\). But this would yield one number for
each pair of communities \((510 × 509 / 2)\): too much data to deal with,
and it is hard to get a sense of the statistical significance of the
resulting correlation coefficient.)
